---
title: "Risk-return relationship"
author: "Rodolfo Vanzini"
date: "6/11/2019"
output: 
  html_document: 
    self_contained: yes
  pdf_document:
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, fig.width = 6.0, fig.asp = 0.75, out.width = "67%", fig.align = "center", fig.show = 'hold', dev = c('png'))
```

I want to investigate the risk-return relationship in a 25,308 mutual funds dataset.

First import the data set
```{r import}
library(tidyverse)
(dt_set <- read_csv("data/funds.csv"))
```

Let's select the colums we want.

```{r select-cols}
(f_return <- dt_set %>% select(fund_name, contains("fund_mean_annual")))
(f_stdev <- dt_set %>% select(fund_name, contains("fund_standard_deviation")))
```

Rename the columns.
```{r rename}
f_return <- f_return %>% 
  rename(`3years` = fund_mean_annual_return_3years) %>% 
  rename(`5years` = fund_mean_annual_return_5years) %>% 
  rename(`10years` = fund_mean_annual_return_10years)
f_stdev <- f_stdev %>% 
  rename(`3years` = fund_standard_deviation_3years) %>% 
  rename(`5years` = fund_standard_deviation_5years) %>% 
  rename(`10years` = fund_standard_deviation_10years)
```

Gather columns to tidy data.
```{r gather}
library(magrittr)
# beware of the %<>% sign
f_return %<>% gather(-fund_name, key = "time", value = "return")
f_stdev %<>% gather(-fund_name, key = "time", value = "stdev")
```

Left join in one tibble.
```{r join}
dt <- left_join(f_return, f_stdev)
```

Plot the data in a scatter point chart. Since there are so many points (25,308) to plot, adding an `alpha` transparency allows to see where density of funds lies. 
```{r pl-rr-1, out.width="100%", fig.width=7.0}
dt %>% ggplot(aes(stdev, return)) + 
  geom_point(alpha = 1/100) + 
  facet_wrap(~time) + 
  stat_smooth(method = "loess")
```

Remove uninteresting data, reorder `time` and replot.
```{r pl-rr-2, out.width="100%", fig.width=7.0}
dt %>% filter(stdev <= 20) %>% 
  mutate(time = factor(time, levels = c("3years", "5years", "10years"))) %>% 
  ggplot(aes(stdev, return)) + 
  geom_hex(bins = 50) + 
  facet_wrap(~time) + 
  stat_smooth(method = "loess", span = 2.0) +
  ylim(c(-2, +4)) + 
  labs(title = "Risk-return relationship for 25,000 mutual funds", 
       subtitle = "Risk-return relationship is positive across different time horizons", 
       caption = "Analysis by R. Vanzini, data source Morningstar", 
       x = "Standard deviation (percent)", 
       y = "Annual return (percent)")

# ggsave(file = "plots/rr.png", width = 7.0, height = 5.25)
```


